---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


Questions for Sudeshna: 

1. Label symmetry - baseline model is per patient, Tanish is per sequence
2. Tanish has patients in both train and test set 
3. either we need to make structured data sequence level OR make tanish's sequence date patient level 
4. rule: if ANY sequence is labeled as cognitive impairment, do we give patient that label? 
5. How do we handle folding in neither/no? 
6. If we want to employ this strategy, all patients that get a label of cognitive impairment should instantly have a patient label of cognitive concern 


Okay to leave patients in test/train but make sure labels are on patient level 

Ask how long it will take to retrain, ensuring accurate test/train split 

next monday APOE (and maybe COVID NAT)

```{r}
library(glmnet)
library(ROCR)
library(tidyverse)
library(data.table)
library(lubridate)

# read in EDW data 
meds<-fread("data/medications.csv")
diagnoses <- fread('data/icd_diagnosis.csv')
diagnoses[,Date := ymd(ContactDTS)]
meds[,Date := ymd(OrderDTS)]

# date time filters
#diagnoses <- diagnoses[Date >= '2018-01-01' & Date <= '2018-12-31']
#meds <- meds[Medication_Date >= '2018-01-01' & Date <= '2018-12-31']

AD_Meds <- c("Galantamine", "Donepezil", "Rivastigmine", "Razadyne", "Razadyne ER", "Aricept", "Exelon","Namenda","Memantine");

meds[,AD := grepl(paste(AD_Meds, collapse="|"),MedicationDSC,ignore.case = T)]

diagnoses[,AD := (grepl("Dementia|Alzheimer's|cognitive impairment|memory loss|frontotemporal dementia",DiagnosisNM,ignore.case = TRUE) | grepl("^290|^294|^331|G30|G31|780.93",ICD9CD,ignore.case = TRUE) | grepl("^290|^294|^331|G30|G31|780.93",ICD10CD,ignore.case = TRUE)) & !grepl("hydrocephalus",DiagnosisNM,ignore.case = TRUE)];

structured_variables <- data.table(PatientID = unique(c(diagnoses$PatientID,meds$PatientID)))

structured_variables[diagnoses[AD==1,.N,by=.(PatientID)],on=.(PatientID),Dx_Count := i.N]
structured_variables[meds[AD==1,.N,by=.(PatientID)],on=.(PatientID),Med_Count := i.N]
structured_variables[is.na(structured_variables)] <- 0
structured_variables[,AD_Med_or_ICD_Code := ifelse(Dx_Count > 0 | Med_Count > 0,1,0)]


# tanish set 

tanish <- fread('data/sample_8_12.csv')
# need EMPI for tanish's set, get from EDW data
tanish[rbindlist(list(meds,diagnoses),fill =T),on=.(patient_id = PatientID),EMPI := i.EMPI]
# 0 = no, 1 = neither, 2 = CI
tanish[,Binary_CI := ifelse(predictions == 2,1,0)]
# at least one CI, give 1 for patient level cognitive impairment 
tanish[,patient_CI := ifelse(sum(Binary_CI) > 0,1,0),by=.(patient_id)]
tanish[,num_sequences_patient := .N,by=.(patient_id)]

# one row per patient 
tanish_predictions <- tanish[,.SD[1],by=.(patient_id)][,.(patient_id,patient_CI,num_sequences_patient,EMPI)]
# merge structured data back on 
tanish_predictions[structured_variables,on=.(patient_id=PatientID),`:=` (Dx_Count = i.Dx_Count, Med_Count = i.Med_Count, AD_Med_or_ICD_Code = i.AD_Med_or_ICD_Code)]

# any missing data? NO
nrow(tanish_predictions[is.na(AD_Med_or_ICD_Code)])

# patient cognitive impairment by whether they have med/code 
tanish_predictions[,.N,by=.(patient_CI,AD_Med_or_ICD_Code)]

# read in APOE data 
apoe <- fread('data/apoe/Partners_biobank_APOE_nodup.csv')
bib <- rbindlist(list(fread('data/apoe/sd587_03212118253982866_6375194824947826181_Bib.txt'),fread('data/apoe/sd587_03212118253982866_6375194824947826182_Bib.txt')))
# join back EMPI from mapping
apoe[bib,on=.(subject_id = Subject_Id),EMPI := i.EMPI]

tanish_predictions[apoe,on=.(EMPI),APOE := i.APOE]

##### need collapsed APOE definition

tanish_predictions[,.N,by=.(patient_CI,AD_Med_or_ICD_Code,APOE)]

fwrite(tanish_predictions,file='data/tanish_predictions_with_structured_features.csv')

# view false negatives
View(tanish[patient_id %in% tanish_predictions[patient_CI == 0 & AD_Med_or_ICD_Code == 1,patient_id]])
#view false positives
View(tanish[patient_id %in% tanish_predictions[patient_CI == 1 & AD_Med_or_ICD_Code == 0,patient_id]])
fwrite(tanish[patient_id %in% tanish_predictions[patient_CI == 1 & AD_Med_or_ICD_Code == 0,patient_id]][order(patient_id,-predictions)],file='data/false_positive_sequences.csv')





##### TANISH - ACCURACY MAXIMIZING THRESHOLD #####

library(ROCR)
library(data.table)

# put your test set with binary probs here, 1 col named 'probability' and one named 'annotator_label'
tanish_test_set <- fread("data/test_binary_classification.csv")
# tanish_test_set[,annotator_label := ifelse(annotator_label == 2,1,0)]
# tanish_test_set[annotator_label == 0,probability := 1 - probability]

acc <- performance(prediction(tanish_test_set$probability,tanish_test_set$annotator_label),measure = 'acc')
acc <- data.table(cutoffs = acc@x.values[[1]],accuracy = acc@y.values[[1]])
threshold <- acc[acc[,.I[which.max(accuracy)]],cutoffs]

# .729 sens
sens <- performance(prediction(tanish_test_set$probability,tanish_test_set$annotator_label),measure = 'sens')
sens <- data.table(cutoffs = sens@x.values[[1]],sens = sens@y.values[[1]])
sens[cutoffs == threshold]

# 0.9346939	
spec <- performance(prediction(tanish_test_set$probability,tanish_test_set$annotator_label),measure = 'spec')
spec <- data.table(cutoffs = spec@x.values[[1]],spec= spec@y.values[[1]])
spec[cutoffs == threshold]

# 0.9343537
auc <- performance(prediction(tanish_test_set$probability,tanish_test_set$annotator_label),measure = 'auc')

## read in gold standards 
# gold <- fread("data/Dementia_ReferenceStandardDataset_08292019_subset_mrn.csv")
# gold[,Label := ifelse(syndromic_dx %in% c(1,2,3,4,9),1,0)]
# structured_variables[gold,on= .(EMPI = empi),Label := i.syndromic_dx]
# 
# test_set_EMPIs <- fread('test_empi_0.csv')
# test_set <- structured_variables[EMPI %in% test_set_EMPIs$EMPI]
# structured_variables <- structured_variables[!EMPI %in% test_set_EMPIs$EMPI] 
# train_set_EMPIs <- fread('train_empi_0.csv')
# structured_variables <- structured_variables[EMPI %in% train_set_EMPIs$EMPI] 
# 
# structured_variables[,.N,by = .(Label)]
# structured_variables[,EMPI := NULL]
# test_set <- test_set[,.(Dx_Count,Med_Count,Label,EMPI)]
# structured_variables[,Label := ifelse(Label %in% c(1,2,3,4,9),1,0)]
# test_set[,Label := ifelse(Label %in% c(1,2,3,4,9),1,0)]




```

Modeling: ONLY RUN IF BASELINE MODEL MAKES SENSE 
```{r}
set.seed(5281995)
f <- as.formula(Label ~ .)
x <- model.matrix(f, structured_variables)[, -1]
y <- as.matrix(structured_variables[,3])

fit = cv.glmnet(x, y, family = "binomial",type.measure = "auc")

plot(fit)
coef(fit, s = "lambda.min")

saveRDS(fit,file = "glmnet.rds")

predict(fit, newx = x, s = "lambda.min", type = "class")

results <- data.table(Prediction = predict(fit, newx = x, s = "lambda.min", type = "class")[,1],Target = y)

results[,Correct := ifelse(Prediction == Target.Label,1,0)]
results[,sum(Correct)]  


# need new x 

#AUC_ROC	ACC	MCC	TP	TN	FP	FN	PPV (Precision) 	NPV	TPR (Sensitivity, Recall)	TNR (Secificity, selectivity)	F1-score


perf_auc <- performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'auc')
print(perf_auc@y.values)  # .794

# write out predictions for Ayush 
test_x <- model.matrix(f, test_set)
test_y <- as.matrix(test_set[,3])
roc_data <- data.table(Prediction = predict(fit, newx = test_x, s = "lambda.min", type = "class")[,1],Target = test_y)
roc_data[,Correct := ifelse(Prediction == Target.Label,1,0)]
roc_data[,sum(Correct)]  


fwrite()
```


```{r}

sens_spec <- performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"),  as.matrix(test_set[,3])), "tpr", "fpr")

saveRDS(sens_spec, "sens_spec.RDS")
saveRDS(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"),  as.matrix(test_set[,3])), "predict_prob.RDS")
saveRDS(as.matrix(test_set[,3]), "Test_labels.RDS")

perf_acc <- performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'acc')
print(perf_acc@y.values) # .81

# TP
res <- data.table(preds = predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="class"),truth =  as.matrix(test_set[,3]))
res[,.(TP = preds.1 == '1' & truth.Label == 1)][,.N,by=.(TP)] # 18

# TN
res[,.(TN = preds.1 == '0' & truth.Label == 0)][,.N,by=.(TN)] # 43

# FP
res[,.(FP = preds.1 == '1' & truth.Label == 0)][,.N,by=.(FP)] # 0

# FN
res[,.(FN = preds.1 == '0' & truth.Label == 1)][,.N,by=.(FN)] # 16

metrics <- data.table(
cutoffs = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'ppv')@x.values[[1]], perf_ppv = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'ppv')@y.values[[1]], perf_npv = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'npv')@y.values[[1]], perf_tpr = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'tpr')@y.values[[1]], perf_tnr = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'tnr')@y.values[[1]], perf_f1 = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'f')@y.values[[1]], acc = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'acc')@y.values[[1]])


# do 1 minus this so mirror image
jpeg("ROC.jpeg",height=500,width=500)
plot(sens_spec,cex.axis=5)
dev.off()

fpr_tpr <-performance(prediction(predict(fit, s=fit$lambda.min, x, type="response"), y), "tpr","fpr")

cutoffs <- data.table(cut=fpr_tpr@alpha.values[[1]], fpr=fpr_tpr@x.values[[1]], 
                      tpr=fpr_tpr@y.values[[1]])

cutoffs <- cutoffs[order(-tpr)]   # 0.6402498 seems to be optimal threshold to get sensitivity equal to current test 

fwrite(cutoffs,file="sens_spec_cutoffs.csv")


new_results[,.N,by=.(Target.Label)] # 452 true negatives, 452/778 = .58097


best.threshold <- .35
ux <- predict(fit,as.matrix(test_set[,1:2]),s=fit$lambda.min, type="response")
probs.new <- ifelse(ux>best.threshold,1,0)
new_results <- data.table(Predicted = probs.new[,1],Target = as.matrix(test_set[,3]))
new_results[,Correct := ifelse(Predicted == Target.Label,1,0)]
new_results[,sum(Correct)] 

new_results[,EMPI := test_set$EMPI]
new_results[gold,on=.(EMPI = empi),Label := i.syndromic_dx]

fwrite(new_results,file= 'results_structured_only.csv')

# best.threshold <- 0
# ux <- predict(fit,x,s=fit$lambda.min, type="response")
# probs.new <- ifelse(ux>best.threshold,1,0)
# new_results <- data.table(Predicted = probs.new[,1],Target = y)

# new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0])
# 
# # values between  0.3983138 and .4001
# 
# best.threshold <- .3983138
# ux <- predict(fit,x,s=fit$lambda.min, type="response")
# probs.new <- ifelse(ux>best.threshold,1,0)
# new_results <- data.table(Predicted = probs.new[,1],Target = y)
# 
# new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0])



threshold_function <- function(threshold){
  best.threshold <- threshold;
  ux <- predict(fit,x,s=fit$lambda.min, type="response");
  probs.new <- ifelse(ux>best.threshold,1,0);
  new_results <- data.table(Predicted = probs.new[,1],Target = y);
  new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0]);
}

test_fpr <- map_dfr(c(seq(.3983138,.4001,by=.00001),.42),~data.table(Cut = .x,FPR = threshold_function(.x)))

threshold_random <- function(threshold){
  best.threshold <- threshold;
  ux <- predict(fit,x,s=fit$lambda.min, type="response");
  ux <- ux + rnorm(length(ux),.00001,.0001);
  probs.new <- ifelse(ux>best.threshold,1,0);
  new_results <- data.table(Predicted = probs.new[,1],Target = y);
  new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0]);
}


# number of rows that are equal to min response 

responses <- data.table(Response = predict(fit,x,s=fit$lambda.min, type="response"))
responses <- responses + rnorm(nrow(responses),.0000001,.000001);

perf <- performance(prediction(responses, y),'auc')
print(perf@y.values)
sens_spec <- performance(prediction(responses, y), "tpr", "fpr")

plot(sens_spec,cex.axis=5)

```
